---
title: "GOVT6139</br>Research design"
subtitle: "Week 04</br>Review of the Literature"
author: "Francesco Bailo"
institute: "The University of Sydney"
date: "Semester 1, 2024 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling

---

background-image: url(https://upload.wikimedia.org/wikipedia/en/6/6a/Logo_of_the_University_of_Sydney.svg)
background-size: 95%

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = 'svg', out.width = "45%", fig.width = 6,
                      fig.align="center")

```

---

## Acknowledgement of Country

I would like to acknowledge the Traditional Owners of Australia and  recognise their continuing connection to land, water and culture. The  University of Sydney is located on the land of the Gadigal people  of the Eora Nation. I pay my respects to their Elders, past and present.

---

class: inverse, center, middle

# Chapter 3

# The Uses of Theory

.center[Slides adapted from Creswell, Research Design 6e, SAGE Publishing, 2023 for GOVT6139 Research Design.

Do not reshare]

---

class: middle, center

# Any question concerning the content for this week?

---

## Today's class

TBC

---

# Chapter 3 Learning Objectives

1. Identify the seven steps in locating and using theories for incorporation into a proposal or a study.

2. Identify the major distinctions between an inductive and deductive use of theory in a research study.

3. Define variables, their measurement, and a causal model for inclusion in quantitative research.

4. Distinguish between the deductive approach, a theoretical standpoint, and an end point as ways to incorporate theories into qualitative research.

5. Describe the difference between discipline-based theory and social justice use of theories in mixed methods research.

---

# Introduction

What theories might be used?

Quantitative researchers test hypotheses from theory

Qualitative researchers may use as a lens or generate a theory

Mixed methods researchers may do both


---

## Two types of reasoning: Induction and deduction

### Relationship between observations (evidence) and theory

.center[<img src = '../img/from-theory-to-hypothesis-a.png' width = '70%'></img>]

---

# The Varied Use of Theory

1. Identify the theory from literature search

2. Place the theory at the beginning of the article as an *a priori* framework

3. Name the theory to be used and describe how it informs the quantitative, qualitative or mixed methods study

4. Discuss studies using the theory

5. Diagram the theory

6. Theory provides framework for the study

7. Review theory’s role at end of study in findings and results, and compare to other studies

---

# Definition of a Theory in Quantitative Research

* Theory is a scientific prediction or explanation of what the Researcher expects to find

* Interrelated set of constructs or variables

* Hypothesis: predictive statement that is tested

* Theory emerges from testing hypotheses over and over

The independent variable X has an effect on the dependent variable Y;

<center>

```{r dia1, echo = FALSE, out.height='80%', out.width = '80%', fig.align='center'}

require(DiagrammeR)

grViz("                           
digraph diagram {   
  
  # graph statement
  #################
graph [rankdir = LR]
node [shape = oval, style = filled, fillcolor = Linen]
  
  'Independent variable X'
  'Dependent variable Y'
  
  # edges
  #######
  'Independent variable X'   -> 'Dependent variable Y' [label = 'affects']
}
")

```

</center>

Then a theory would explain why...

---

# A theory in quantitative research can explains phenomena at different levels 


__Micro-level theories__  – limited to a small segment of time, place or number of people

__Meso__  __-level theories__  – link micro-level and macro-level theories

__Macro-level theories__  – explain larger scales, for example the entire society

---

# Variables in Quantitative Research

__Variable__  – characteristic of an individual or group that can be **measured** or **observed**

Examples: age, gender, socioeconomic status, attitudes, but also more abstract characteristics such the quality of democracy in a country, 

Variables may have temporal order or be measurable or observable

* Independent variable (or predictor variable)
  * Variables that probably cause or affect outcomes
  
* Dependent variable (or outcome variable)
  * The outcomes that depend on the independent variables
  
---

# Variables in Quantitative Research

* Intervening or mediating
  * Variables that stand between the independent and dependent variables
  * Transmit the effects
  
* Moderating
  * Predictor variables that *affect* the relationship between independent and predictor variables
  
* Confounding
  * “Third variable” that is both related to the independent and dependent variables

---

## Dependent and independent variables

<center>

```{r dia2, echo = FALSE, out.width = '100%', fig.align='center'}

require(DiagrammeR)

grViz("                           
digraph diagram2 {   
  
  # graph statement
  #################
graph [rankdir = LR]
node [shape = oval, style = filled, fillcolor = Linen]
  
  'Independent variable X'
  'Dependent variable Y'
  
  # edges
  #######
  'Independent variable X'   -> 'Dependent variable Y' [label = 'affects']
}
")

```

</center>

---

## Dependent, independent and mediating variables

<center>

```{r dia3, echo = FALSE, out.width = '100%', fig.align='center'}

require(DiagrammeR)

grViz("                           
digraph diagram3 {   
  
  # graph statement
  #################
graph [rankdir = LR]
node [shape = oval, style = filled, fillcolor = Linen]
  
  'Independent variable X'
  'Dependent variable Y'
  'Mediating variable Z'
  
  # edges
  #######
  'Independent variable X' ->  'Mediating variable Z' [label = 'affects']
  'Mediating variable Z' -> 'Dependent variable Y' [label = 'affects']
}
")

```

</center>


---

## Dependent, independent and moderating variables

<center>

```{r dia4, echo = FALSE, out.width = '100%', fig.align='center'}

require(DiagrammeR)

grViz("                           
digraph diagram3 {   
  
  # graph statement
  #################
graph [rankdir = LR]
node [shape = oval, style = filled, fillcolor = Linen]
  
  'Moderating variable Z'
  'Independent variable X'
  'Dependent variable Y'
  'effect' [shape = circle, fillcolor = Orange]
  
  # edges
  #######
  'Moderating variable Z' -> 'effect' [label = 'affects']
  'Independent variable X' ->  'effect' [arrowhead = none]
  'effect' -> 'Dependent variable Y'
  
  
}
")

```

</center>


---

## Dependent, independent and confounding variables

<center>

```{r dia5, echo = FALSE, out.width = '100%', fig.align='center'}

require(DiagrammeR)

grViz("                           
digraph diagram3 {   
  
  # graph statement
  #################
graph [rankdir = LR]
node [shape = oval, style = filled, fillcolor = Linen]
  
  'Confounding variable Z'
  'Independent variable X'
  'Dependent variable Y'
  
  # edges
  #######
  'Confounding variable Z' -> 'Independent variable X' [label = 'affects']
  'Confounding variable Z' -> 'Dependent variable Y' [label = 'affects']
  'Independent variable X' -> 'Dependent variable Y' [label = 'affects']
  
  
}
")

```

</center>


---

# Hupothesis and Scale in quantitative research

  * __Hypothesis __ – prediction about a specific event or relationship between variables
  * __Scale of Measurement__  – response options available to participants on instruments or collected by researcher observing participants
    * __Categorical scales __ – nominal and ordinal
    * __Continuous scales __ – interval and ratio

---

# Stating Theories in Quantitative Research

## Theory

A theory should provide an explanation of why something happens

  * X affects Y.

## Hypothesis 

An hypothesis should provide a statement that is testable if we measure both X and Y

  * **If** X is observed **then** Y is also observed.  

---

# Placement of Quantitative Theories

* Towards the beginning of the study
* Used deductivel
* Framework for study
* Organizing model for
  * Research questions or hypotheses
  * Data collection procedure

---

# Variation in Theory Use in Qualitative Research

1. Used for a **broad explanation**

2. Used a  __theoretical lens or perspective or standpoint__  – overall orienting lens for question of gender, class, and race

3. Theory becomes the end point in an **inductive approach**

4. Do not employ any explicit theory

---

# Qualitative theoretical perspectives

* Feminist perspective
* Racialized discourse
* Critical theory
* Queer theory
* Disability inquiry

---

Figure 3.5:  The **Inductive Logic** of Research in a Qualitative Study (Creswell & Creswell, 2022)

.center[<img src = '../img/Creswellrd6e_PPT035.jpg' width = '35%'></img>]

---

# Locating the Theory in Qualitative Research

## Locating the theory in qualitative research:

* Studies with cultural themes or theoretical lens
  * In the opening passages
* Studies with an emerging design of qualitative inquiry
  * At the beginning, may be modified
* Studies based on grounded theory
  * Theory derived inductively
  * Placed at the end

---

# Types of Mixed Methods Theory Use

* Mixed methods studies may:
  * Include theory deductively (theory testing and validity)
  * Include theory inductively (an emerging theory or pattern)
  * Use a discipline-based or participatory-social justice framework

---

# Discipline-based theory use

* Placed **at the beginning** as an a priori framework to guides questions/hypothesis

* Named and described, used to inform the explanation of the major variables in the study

* Diagrammed to indicate the causal links

* Used to guide the data collection process

* Returned to at the end to inform findings and results

---

# Participatory-social justice theory use

* Growing use of **transformative theory** and mixed methods

* Theoretical framework to inform a mixed methods project

* Ethical stance of inclusion and challenging oppression

* Designed to build trust

* Use of results to enhance social justice

---

# Participatory-social justice theory use

## Incorporating into a mixed methods study

1. Indicate **at the beginning** that transformative theory is in use
2. Early in the study, outline the framework.
3. Connect the theory with the existing body of knowledge on the topic
4. Involve the community of interest in the process of the research
5. Take a stand with the research question
6. Advance in the design qualitative and quantitative methods
7. Talk about your experience as a researcher
8. End the study by advocating for change

---

# Distinctions Between a Theory and a Worldview

* Theories and worldviews are not the same

* Worldview relates to *beliefs* and *values* of researcher and how these inform the study

* Theories are less abstract, drawn from literature, and guide many research process aspects (e.g., methodological approaches to design, data collection methods)

---

Figure 3.6:  The Four Levels for Developing a Research Study (Adapted from Crotty, 1998, by Creswell & Creswell, 2022)

.center[<img src = '../img/Creswellrd6e_PPT036.jpg' width = '40%'></img>]

---

# Summary

Theory is used to:

* Provide an explanation or prediction about the relationship among variables in quantitative research

* Provide broad explanation, a theoretical lens, or as an end-point in qualitative research

* Inform the design, data collection, analysis, and interpretation in mixed methods research

---

