---
title: "GOVT6139</br>Research design"
subtitle: "Week 05</br>Writing Strategies and Ethical Considerations"
author: "Francesco Bailo"
institute: "The University of Sydney"
date: "Semester 1, 2024 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling

---

background-image: url(https://upload.wikimedia.org/wikipedia/en/6/6a/Logo_of_the_University_of_Sydney.svg)
background-size: 95%

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = 'svg', out.width = "45%", fig.width = 6,
                      fig.align="center")

```

---

## Acknowledgement of Country

I would like to acknowledge the Traditional Owners of Australia and  recognise their continuing connection to land, water and culture. The  University of Sydney is located on the land of the Gadigal people  of the Eora Nation. I pay my respects to their Elders, past and present.

---

class: inverse, center, middle

# Chapter 4

# Writing Strategies and Ethical Considerations

.center[Slides adapted from Creswell, Research Design 6e, SAGE Publishing, 2023 for GOVT6139 Research Design.

Do not reshare]

---


## Course outline

.pull-left[

**Week 1**: Introduction

**PART I PRELIMINARY CONSIDERATIONS (Weeks 2-5)**

**Week 2**: The Selection of a Research Approach

**Week 3**: Review of the Literature

**Week 4**: The Use of Theory

**Week 5**: Writing Strategies and Ethical Considerations + Guest lecture w/t Christopher Pepin-Neff 👈

**PART II DESIGNING RESEARCH (Weeks 6-12)**

**Week 6**: The Introduction **no seminar** (public holiday)

⏸️ *Mid Semester break* ⏸️

]

.pull-right[


**Week 7**: The Purpose Statement

**Week 8**: Research Questions and Hypotheses + Guest Lecture w/t Charles Robert Butcher 

**Week 9**: Quantitative Methods

**Week 10**: Quantitative Methods: Data Analysis Lab (Make sure you install R and RStudio)

**Week 11**: Qualitative Methods + Guest lecture w/t Peter Chen

**Week 12**: Mixed Methods Procedures + Guest lecture w/t Assel Mussagulova

**Week 13**: Conclusions


]

---

## Today's class

| Time         | Content                                         |
|--------------|-------------------------------------------------|
| 1:00 - 2:00  | Chris' Guest Lecture                            |
| 2:00 - 2:20  | How to write your proposal                      |
| 2:20 - 2:40  | Check-in <br/><br/> Task 1: Individual (20 min) |
| 2:40 - 3:00  | Ethical issues                                  |

---

# Chapter 4 Learning Objectives

1. Identify the nine key topics to include in a proposal or a research study.

2. Compare the differences in the structure of proposals for quantitative, qualitative, and mixed methods research studies.

3. Evaluate the quality of your writing for conciseness, coherence, and unnecessary words (refer to textbook!).

4. Identify ethical issues and strategies for addressing them before conducting a study and during and after a study.

---

class: inverse, center, middle

# How to write a proposal

---

# Introduction

Have an idea of the general structure of topics

Structure depends on the design

Be aware of good writing practices

Anticipate ethical issues prior to the study

---

# Topics Presented in a Proposal

## Arguments presented in a proposal

* Consider early on in planning the study

  * The major points that need to be addressed
  
  * How they will convey a cohesive picture of the study
  
  * What the core arguments to advance are

---

## Arguments presented in a proposal (nine key topics, or must-have)

1 What do readers need to better understand the topic?

2 What do readers need to know about the topic?

3 What do you propose to study?

4 What is the setting, and who are the people you will study?

5 What methods do you plan to use to collect data?

6 How will you analyze the data?

7 How will you validate the findings?

8 What ethical issues will your study present?

9 What do preliminary results show about the practicability and value of the proposed study?

---

# Check-in

---

class: middle, inverse

# TASK 1 

Individually,

1. Revise four example outlines on the Task sheet;
2. Ask any question about the items in the examples;
2. Write the outline for your own A2 proposal;
3. Share it the class (you can write it directly on Padlet, or you can write it down and share a photo of it).
4. Go through the check-list on the back of the Task sheet using one of the outlines shared on Padlet.

```{r echo = FALSE}
library(countdown)

countdown(minutes = 20, seconds = 00)
```

---

# Writing Strategies

* Write ideas down early in the process rather than talking about them

* Work through several drafts of a proposal rather than trying to polish the first draft

* Do not edit at the early draft stage

  * Start with an outline
  
  * Write a draft and shift and sort ideas
  
  * Finally, edit and polish each sentence
  
**Also, refer to this week chapter of the textbook for more...**

---

class: inverse, center, middle

# Ethical issues

---

# Ethical Issues

* Research involves collecting data from people

* Researchers need to

  1. Protect research participants
  2. Develop trust with research participants
  3. Promote the integrity of research
  4. Guard against misconduct
  5. Cope with new problems that emerge

---

## Prior to beginning the study:

* Consider codes of ethics

* Apply to the institutional review board (in Australia: Human Research Ethics Committees)

* Obtain the necessary permissions to gain access to the site and to study participants

* Select a site without vested interests

* Negotiate authorship for publication

---

## Beginning the study:

* In the research problem:
  * Identify a problem that will benefit individuals being studied
* In the purpose and questions:
  * Convey the purpose and sponsors of the research to participants
* Do not pressure participants into signing consent forms, obtain informed consent from participants
* Respect norms and charters of indigenous cultures
* Assess the burden of research on the participants

---

## Collecting the data:

* Respect the site, and disrupt as little as possible

* Make sure all participants receive the benefits

* Collaborate with participants

* Avoid deceiving participants

* Respect potential power imbalances, consider reciprocity

* Avoid exploitation of participants

* Avoid collecting harmful information

---

## Analyzing the data:

Avoid taking sides or disregarding data

Avoid disclosing only positive results – reflect the statistical tests and diversity of perspectives

* Analyzing the data:
* Respect the privacy of participants:
  * Protect anonymity of participants
  * Remove names from responses
  * Qualitative researchers use aliases and pseudonyms

---

## Reporting, sharing, and storing data:

* Do not falsify authorship, evidence, data, findings or conclusions

* Do not plagiarize

* Avoid disclosing information that would harm participants

* Communicate in clear straightforward, appropriate language

---

## Reporting, sharing, and storing data:

* Share data with others (e.g. stakeholders, participants)

* Keep raw data and other materials for a reasonable period of time

* Do not duplicate or piecemeal publications

* Complete proof of compliance with ethical issues and lack of conflict of interest

* Understand who owns the data

---

class: inverse

# Summary

* How to write a proposal

* Use one of the outlines for a qualitative, quantitative, or mixed methods proposal

* Importance of putting words on paper

* Consider anticipated ethical issues and describe

---

class: inverse, middle, center

# Thank you! 

Next seminar in two weeks!

A1 is due next Monday (25 March 2024): If you are late apply for a simple extension! 


    



